---
title: "Rescaling of cell and genes features"
author: "Benjamin Arthur"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rescaling of cell and genes features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

The purpose of this package is to rescale cell or gene features from a seurat object which has been normalized and RunNMF has being implemented.

The library and object for rescaling can be found in the singlet package.

```{r}
library(singlet)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(cowplot)
library(Seurat)
pbmc3k <- get_pbmc3k_data()
pbmc3k <- pbmc3k %>% NormalizeData() %>% RunNMF(reps = 1)
```

## Human Gene set from GOBP_cellular_response_to_interferon_beta
A set of human genes from Gene Set Enrichment Analysis (GSEA) sourcing from Human_NCBI_Gene_ID was used for the study.

```{r}
#df<- read_tsv('gopb.tsv',col_names = TRUE)
#dff <-df %>%

 # separate_rows(STANDARD_NAME,sep =',')%>%
 # separate_rows(GOBP_CANONICAL_WNT_SIGNALING_PATHWAY, sep = ',')%>%
 # rename("Genes"="GOBP_CANONICAL_WNT_SIGNALING_PATHWAY")%>%
 # select(Genes)
#names(dff)<- NULL
Features <- as.vector(dff[335:644,])
Feat <- as.vector(dff[600:644,])
# Using these features are giving me error, so i implemented the procrutes with some few cells from the pbmc3k dataset.

```



## Recaling of a gene features with different threshold (Lambda)

In this example we used four genes of interest with 0.8 threshold, and we assume that using a threshold which is close to 1, the relationship between v and H would reduce.

From the plots below, the first UMAP visualized the genes features after rescaling and the second UMAP is visualizing the genes features with recaling.
```{r}
set.seed(542)
cell = c("AACCGCCTCTACGAA","AACCTACTGTGAGG","AACCAGTGATACCG","AACACGTGGAAAGT")

pbmc3k <- RescaleNMF(pbmc3k,features=cell,reduction="nmf",lambda=0.8,reduction.name="nmf_rescaled")

model_reference <- RunUMAP(pbmc3k, reduction = "nmf", reduction.name = "umap_nmf", dims = 1:ncol(pbmc3k@reductions$nmf@cell.embeddings))

model_query <- RunUMAP(pbmc3k, reduction = "nmf_rescaled", reduction.name = "umap_nmf_rescaled", dims = 1:ncol(pbmc3k@reductions$nmf_rescaled@cell.embeddings))

plot_grid(DimPlot(model_reference, reduction = "umap_nmf", group.by = "cell_type"), DimPlot(model_query, reduction = "umap_nmf_rescaled", group.by = "cell_type"), ncol = 2)


```





```{r}
model_procrustes <- procrustesNMF(model_reference,model_query)

plot_grid(DimPlot(model_reference, reduction = "umap_nmf", group.by = "cell_type"), DimPlot(model_procrustes, reduction = "umap_nmf_procrustes", group.by = "cell_type"), ncol = 2)
```

## Implementing procrustes analysis

Procrutes function is used to align the umap coordinate of the query object to the reference object for better view and understanding. I applying this function would recalculate the distance between the object in a way to match them on similar coordinates.

## Applying the rescale Function using lambda of 0.2

```{r}
set.seed(542)
set.seed(542)
# Cell.embedding
cell = c("AACCGCCTCTACGAA","AACCTACTGTGAGG","AACCAGTGATACCG","AACACGTGGAAAGT")

pbmc3k <- RescaleNMF(pbmc3k,features=cell,reduction="nmf",lambda=0.2,reduction.name="nmf_rescaled")

model_reference <- RunUMAP(pbmc3k, reduction = "nmf", reduction.name = "umap_nmf", dims = 1:ncol(pbmc3k@reductions$nmf@cell.embeddings))

model_query <- RunUMAP(pbmc3k, reduction = "nmf_rescaled", reduction.name = "umap_nmf_rescaled", dims = 1:ncol(pbmc3k@reductions$nmf_rescaled@cell.embeddings))

plot_grid(DimPlot(model_reference, reduction = "umap_nmf", group.by = "cell_type"), DimPlot(model_query, reduction = "umap_nmf_rescaled", group.by = "cell_type"), ncol = 2)

```

## Implementing procrustes analysis

Applying the procrutes function on another rescaled object of lambda equal to 0.2

```{r}
model_procrustes <- procrustesNMF(model_reference,model_query)

plot_grid(DimPlot(model_reference, reduction = "umap_nmf", group.by = "cell_type"), DimPlot(model_procrustes, reduction = "umap_nmf_procrustes", group.by = "cell_type"), ncol = 2)



```




